---
title: "Primary endpoint analysis"
output: html_document
date: "2022-12-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(readr)
library(epitools)
library(gtsummary)
library(emmeans)
library(GAMLj3)
```

## 



```{r load_data}
sero_outcomes <- read_csv("../data/serology_outcomes.csv")


sero_outcomes$outcome_serology = as.numeric(!is.na(sero_outcomes$outcome))

inclusion_data = readxl::read_excel(path = '../data/Time_events_16_01_2023_REVISED_MM_WS_Latest_FINAL_19_01_23_merged.xlsx') #C:/Academico/Projetos/Cloroquina/Time_events_16_01_2023_REVISED_MM_WS_Latest_FINAL_19_01_23.xlsx'


outcome_dat = merge(inclusion_data[, c('subjectno','treatment','pcr_event', 'age_calyr', 'sex', "anychronic", "comorbidities", "diabetes", "bmi", "PCRcovid", "Serumcovid", "DBScovid" )], 
                    sero_outcomes[,c('subjectno','outcome_serology',
                                     'serum_available','dbs_available',
                                     'type_data')],
                    by = 'subjectno', all.x=T)
symptomatic = read_delim('../data/Symptomatic.csv',delim = ';')
outcome_dat = merge(outcome_dat, symptomatic, by='subjectno')


```



# make endpoints

```{r, echo=F}
outcome_dat$Primary_endpoint = (outcome_dat$outcome_serology==1 | outcome_dat$pcr_event==1) & outcome_dat$symptomatic=='Symptomatic'
writeLines('Primary endpoint (symptomatic COVID-19):')
table(outcome_dat$Primary_endpoint, outcome_dat$treatment)

outcome_dat$Asymptomatic_SARSCoV2 = (outcome_dat$outcome_serology==1 | outcome_dat$pcr_event==1) & outcome_dat$symptomatic=='Asymptomatic'

writeLines('Asymptomatic SARS-CoV-2:')
table(outcome_dat$Asymptomatic_SARSCoV2, outcome_dat$treatment)

writeLines('All PCR confirmed Symptomatic COVID-19:')
table(outcome_dat$pcr_event, outcome_dat$treatment)

writeLines('All serologically confirmed SARS-CoV-2 infections:')
outcome_dat$Serology_COVID19 = outcome_dat$outcome_serology==1 & 
  outcome_dat$symptomatic=='Symptomatic'
table(outcome_dat$Serology_COVID19, outcome_dat$treatment)

writeLines('All Symptomatic COVID-19 serologically confirmed using serum:')

outcome_dat$Serum_COVID19 = outcome_dat$outcome_serology==1 &
  outcome_dat$serum_available==1 &
  outcome_dat$symptomatic=='Symptomatic'
table(outcome_dat$Serum_COVID19, outcome_dat$treatment)

writeLines('All Symptomatic COVID-19 serologically confirmed using DBS:')
outcome_dat$DBS_COVID19 = outcome_dat$outcome_serology==1 &
  outcome_dat$dbs_available==1 & outcome_dat$serum_available==0 &
  outcome_dat$symptomatic=='Symptomatic'
table(outcome_dat$DBS_COVID19, outcome_dat$treatment)
getwd()

# Transform treatment variable to facotr and change reference level to placebo
outcome_dat$treatment <- as.factor(outcome_dat$treatment)
outcome_dat$treatment <- relevel(outcome_dat$treatment, ref = "Placebo")

#write_csv(x = outcome_dat,file = '../data/outcomes_derived.csv') File already in the repository
```

# Risk ratio estimation for primary endpoint


```{r}
riskratio(table(outcome_dat$Primary_endpoint, outcome_dat$treatment),rev='rows',method = 'wald')

```


# Recreate Fisher's Exact Test p-value

```{r}
# Criar uma tabela de contingência
tabela <- table(outcome_dat$Primary_endpoint, outcome_dat$treatment)
fisher.test(tabela)

```



# Create models and calculate p-value

## Primary endpoint (All-Cause PCR-Confirmed respiratory illness)

### Treatment as covariable
```{r}

model_1 <- glm(Primary_endpoint ~ treatment, data = outcome_dat, family=poisson(link='log'))
tbl_regression(model_1, exponentiate = T)
```


### Treatment and age as covariable
```{r}

outcome_dat$sex <- as.factor(outcome_dat$sex)
outcome_dat$Primary_endpoint <- as.integer(outcome_dat$Primary_endpoint)


# All-Cause PCR-Confirmed respiratory illness adjusted by age and sex
GAMLj3::gamljglm(
    formula = Primary_endpoint ~ treatment + sex + age_calyr + treatment:sex,
    data = outcome_dat,
    posthoc = ~ treatment,
    model_type='poisson')

```

### Mean a median age
```{r}
mean(outcome_dat$age_calyr)
median(outcome_dat$age_calyr)
```

## Primary endpoint (PCR-confirmed COVID-19)

### Treatment as covariable

```{r}
# Results reporter in the orginal paper and used in the meta-analysis

model_3 <- glm(pcr_event ~ treatment, data = outcome_dat, family=poisson(link='log'))
tbl_regression(model_3, exponentiate = T)

```

### Treatment and age as covariable


```{r}
outcome_dat$pcr_event <- as.integer(outcome_dat$pcr_event)
outcome_dat$treatment <- relevel(outcome_dat$treatment, ref = "Chloroquine/Hydroxychloroquine")

# PCR-confirmed COVID-19 adjusted by age and sex

GAMLj3::gamljglm(
    formula = pcr_event ~ treatment + sex + age_calyr + treatment:sex,
    data = outcome_dat,
    posthoc = ~ treatment,
    model_type='poisson')

```



